Improved state-space models for inference about spatial and temporal variation in abundance from count data

### 2014-11-14 response-to-revision-comments.tex

The authors check for the reliability of their model to recover their parameters by using a 
simulation study. The simulation study covers a large range of different population 
dynamics as well as additional cases where the simulated data deviates from the assumption 
of the model used for the analysis. In addition, to the simulation study, the models are 
applied to real data.

\vspace{0.5cm}
\textit{Thank you for your positive comments!}
\vspace{0.5cm}

In a general way, effort should be put on tightening the text and better focus
on the main goal of the paper, that is about including models of population
growth in a framework that account for imperfect detection of individuals. The
choice of the model of population growth might be subjective and/or motivated by
the knowledge of the life history of the species under study. Even thought, some
studies, proposed some procedures based on model selection to discriminate among
the different models of population growth. Hence, the Ms would benefit of a
better description of the underlying ecological assumptions of the different
models of population growth. For example, the exponential model describes
population growth in an unlimited environment, where $r$ is the maximum per capita
of increase, etc… . In Ricker model, the density-dependence becomes stronger at
higher densities, due to the exponential function relating to carrying capacity.
The Gompertz model is similar to the Ricker model except that there is a
constant linear decrease in the instantaneous growth rate ($r$) as the natural
logarithm of population increase, etc…..

\vspace{0.5cm}
\textit{We have described the classical population growth models in more 
detail (lines ).}
\vspace{0.5cm}

@@

I find that basing the introduction upon the notion of state-based model is rather confusing
regarding the real goal of the paper (see above). The Dail and Madsen model and state-space
models are defined in two different ``statistical worlds''. On the one hand, the constant
Dail and Madsen model (2011) has a Markov property so the variables can only take on a
finite set of discrete values for a finite period of observation. The distinguish feature of
Markov chains is that the conditional probability of $x_{t}$ given the entire past depends only on
the immediately preceding variable $x_{t-1}$. In addition, the errors of observation are well
identified because they are related to imperfect detection of individuals and expressed with
a discrete probability distribution. On the other hand, the state-space model framework
provides a structure for extending time-series models to handle both observation and
process errors. The term ``state-space'' model refers to models where the process model has
an autoregressive property and where the errors have a Gaussian distribution.
Autoregressive model are defined for continuous variable, and so the errors are normally
distributed. A space-state model, being given its properties, can be estimated by using a
Kalman filter. The Kalman filter is an updating procedure whereby the prior estimate of the
state is corrected by how well the next observation can be predicted.

Even if the Dail and Madsen model also provides a structure to account for time dependence
as well as for ``process errors'' through the birth and death process and for imperfect
detection, the Dail and Madsen model and state-space models are defined in two different
``statistical worlds''. Hence, the analytical results of the Dail and Madsen model and a state-space
model would be different. That is why, in my point view, it is unclear and misleading to
relate the Dail and Madsen model to a state-space model.

\vspace{0.5cm}
\textit{Our definition of a state-space model is in agreement with the brand new
book on the subject \citep{newman_etal:2014}. They cover many state-space 
models that have
a first-order Markov state process and a non-Gaussian
observation process. For this reason, we think that our title connects
our work to the bigger picture, which is state-space
models of count data.} 
\vspace{0.5cm}

@@

Lines 118 - 174: the description of the model could be shortened (or moved in part in the
appendices) because it has been described now in further publications.

\vspace{0.5cm}
\textit{We have shortened the DM model description section somewhat.}
\vspace{0.5cm}

@@

Line 135: typography error in \textit{``The third process describes the relationships between''} should
be \textit{``The third process describes the relationships between.''}

\vspace{0.5cm}
\textit{We have fixed the typographic error (surplus ``the'') here.}
\vspace{0.5cm}

Line 190: More explanations on the Gompertz model should be added here, especially on
the link between the Ricker model and the Gompertz model. The original reference of the
Gompertz logistic growth equation should be added here.

\vspace{0.5cm}
\textit{We have described the Gompertz model in more 
detail (lines ).}

@@

Lines 281 - 296: The paragraph about the simulation of the data needs to be clarified to
understand better how the data are simulated. Some references to the different equations
of the model should help to clarify this section. For the simulation study, the time series
which are simulated are rather long (40 years) and also the length of the time series of the
application are about 40 years. If the time series were shorter, how the estimation of the
parameters would be affected.

\vspace{0.5cm}
\textit{We have added clarification and equation references to this paragraph. We expect that
the models would not perform as well with far fewer years of data, and have commented on
this in the discussion (lines).}
\vspace{0.5cm}

Discussion: The text of the discussion needs to be reduced and tightened. The authors
should better focus on discussing the results of this study and relate them to other studies
on models of population growth.

\vspace{0.5cm}
\textit{We have reduced the text of the discussion to focus more on the results of this
study and other studies of population growth.}

@@

\vspace{0.5cm}

The figure A1 and A2 describe results of the simulation based on the Ricker and Gompertz
models, and both are much more ``realistic'' than the model of exponential growth when
applied to population ecology. So, the Fig. A1 and A2 represent core results of the
manuscript and should be rather presented in the main text than in appendices.

\vspace{0.5cm}
\textit{We have put the exponential, Ricker, and Gompertz simulation results tables in the 
main text, and moved the exponential simulation results figure to the appendix.}
\vspace{0.5cm}

@@

Lines 424-429: Could difficulty in estimating vital rates derives from having
survival and recruitment parameters taking similar values? For example many
small songbirds have apparent survival rates of $\sim .5$; for a stable population,
recruitment would be similar, which in lieu of direct information on vital
rates, might make the 2 parameters difficult to separate. In the American Robin
example in Dail and Madsen (2011), survival and recruitment were estimated at
reasonable values of .7 and .5 (although precision was still fairly low). Could
low detection probability have something to do with this problem? From the p
estimates in Fig. B1, detection probabilities were $< .2$ (which seems low to me
given that ovenbirds are loud and sing a lot, although I suppose it has to do
with large sample radius). In the simulations in the Dail-Madsen paper, lowest
coverage for gamma and omega was found at lowest p values (p = .25). I realize
I'm stretching a bit here, but it was a little
unsatisfying to me that the vital rates parameters could not be estimated, and
the reason for why this was so wasn't too clear. The authors suggest individual
heterogeneity might be an issue and suggest incorporating prior information or
data via integrated population models to improve vital rate estimation. However,
I wonder if some additional conjecture might be added here?

\vspace{0.5cm}
\textit{We have added some speculation in this direction (lines ).}
\vspace{0.5cm}
